### 2025-05-24 ms.qmd
updates

```{r}
#| label: setup

pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)

source(here("R", "func.R"))
```

@@

### CORT measurement and validation

@@

## Results

```{r, dataload}
#| label: dataload
#| echo: false
#| warning: false

# Load data
	data  <-  read.csv("./data/Stressor.csv")
```

```{r, dataexplore}
#| label: dataexplore
#| echo: false
#| warning: false

# Divide Treatment in CORT and Temperature Transform mass into mg
data <- data %>%
      mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
      mutate(mass=mass*1000) %>%
data.frame()
data$cort <- factor(data$cort, 
        levels = c("A" ,"B"),
        labels = c("Control", "CORT"))
data$temp <- factor(data$temp,
        levels = c("23", "28"),
        labels = c("Cold", "Hot"))
data$Day <- factor(data$Day,
        levels = c("Before", "After"))

# MASS
# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
 names_from = Day, values_from = mass)

# Estimate mass increase and rescale it to avoid negative values
data_2 <- data_2 %>% 
 mutate(delta_mass=After-Before) %>%
 mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
  NA, 
  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
write.csv(data_2, file = "./data/Stressor_2.csv")

# Split new database by species
deli_stressor_2 <- data_2 %>% filter (species == "delicata")
guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")

# Explore distribution of mass increase
  # A) L.deli
plot_mass <- ggplot(deli_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "temp ~ cort") +
  theme_minimal() +
  facet_grid(as.factor(deli_stressor_2$cort)~as.factor(deli_stressor_2$temp), scales = "free_y")
print(plot_mass)

mass_normal_test<-by(deli_stressor_2$rescaled_delta_mass, interaction (deli_stressor_2$cort, deli_stressor_2$temp), function(x) shapiro.test((x)))
print(mass_normal_test)

  #B) L.guich
plot_mass <- ggplot(guich_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(guich_stressor_2$temp),as.factor(guich_stressor_2$cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(guich_stressor_2$cort~guich_stressor_2$temp, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(guich_stressor_2$rescaled_delta_mass, interaction (guich_stressor_2$cort, guich_stressor_2$temp), function(x) shapiro.test((x)))
print(mass_normal_test)

# BEHAVIOUR
# Split data by species in original database
deli_stressor <- data %>% filter (species == "delicata")
guich_stressor <- data %>% filter (species == "guichenoti")
# Explore distribution of latency to move/time frozen
  # A) L. deli
plot_freeze <- ggplot(deli_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_freeze)

freeze_normal_test<-by(deli_stressor$lat_to_move, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test(log(x+1)))
print(freeze_normal_test)

 #B) L.guich
plot_freeze <- ggplot(guich_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_freeze)

freeze_normal_test<-by(guich_stressor$lat_to_move, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test(log(x+1)))
print(freeze_normal_test)

```{r, models}
#| label: models
#| echo: false
#| warning: false

@@

# Plot time frozen L. delicata
plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 450) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
  ) +
  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_deli_BEH)
ggsave("./output/figures/fig_deli_BEH.png", plot=plot_deli_BEH, width = 15, height = 12, units = "cm", dpi = 3000)

# Plot time frozen L. guichenoti
plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 450) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
  ) +
  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_guich_BEH)
ggsave("./output/figures/fig_guich_BEH.png", plot=plot_guich_BEH, width = 15, height = 12, units = "cm", dpi = 3000)

>>>>

```{r}
#| label: setup
pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
```

```{r, cleandata}
#| label: cleandata
# Obtain the main df using "./R/1_data_process.R"
source(here("R", "data_process.R"))
```

```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the main database.
source(here("R", "func.R"))
#
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(k in 1:length(hormone)){
  for(l in 1:length(temperature)){
    list_name <- paste0(hormone[k], "_", temperature[l])
    n_list[[list_name]] <- sample(df = mass_clean, corti = hormone[k], therm = temperature[l])
  }
}
#
```

```{r, countclutches}
#| label: countclutches
# Count the number of clutches per species
#
clutches <- mass_clean %>% 
  distinct(clutch) %>% 
  nrow()
```

@@

@@

## Results
```{r, models_mass}
#| label: models
#| echo: false
#| warning: false

@@

# Plot time frozen L. delicata
plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 10) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
  ) +
  labs(y = "Time frozen (log-transformed)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_deli_BEH)
ggsave("./output/figures/fig_deli_BEH.png", plot=plot_deli_BEH, width = 15, height = 12, units = "cm", dpi = 3000)

# Plot time frozen L. guichenoti
plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = log(lat_to_move+1), fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 8) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
  ) +
  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_guich_BEH)
ggsave("./output/figures/fig_guich_BEH.png", plot=plot_guich_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
